---
title: "Recipes"
author: "Xuhao Dong"
date: "`r format(Sys.time(), '%d %B, %Y')`"
html_document:
    toc: true
    toc_depth: 6
    number_sections: true
    toc_float: true
    code_folding: hide
    theme: flatly
    code_download: true
---

```{r setup, include=FALSE}
# Just turing off the conflicts because I'm tired of seeing them.
library(tidyverse,warn.conflicts = F)
library(tidymodels,warn.conflicts = F)
library(lubridate,warn.conflicts = F)
library(timeDate)
knitr::opts_chunk$set(echo = TRUE)
```


## Subaru Data Recipes

This assignment asks you to do a few discrete tasks using `recipes` to 
manipulate data in R. We'll work with two data sets: 

* `subaru_data.txt`: We discussed this data in class this week. This data set has
27K rows of Craigslist car listings. The response variable we'll ultimately care
about is `price`. There are a mix of helpful and unhelpful columns available 
to us. 

Additionally, I have a task for you to perform that requires you to work 
with some time series information in R using `lubridate`. 

```{r message=F}
d.subaru <- read_tsv("subaru_data.txt") %>% 
  mutate(odometer = as.numeric(odometer),
         year = as.numeric(year))

d.dates <- tibble(dt = seq(ymd("2018-01-01"),
                             ymd("2020-12-31"),
                             by="days")) %>% 
  mutate(yr = year(dt),
         mon = month(dt,label=T),
         day = day(dt),
         week_day = wday(dt,label=T))

d.subaru <- d.subaru %>%
  mutate(price = as.numeric(price),
    model = as.factor(model),
    condition = as.factor(condition))

d.subaru$year <- as.integer(d.subaru$year)
is.na(d.subaru$year) <- d.subaru$year > 2017

## Funny thing here, I was trying to figure out the max odomemter read. And I looked up the Guninness World Record is acutally 3,000,000 miles. 
d.subaru$odometer <- as.integer(d.subaru$odometer)
is.na(d.subaru$odometer) <- d.subaru$odometer > 300000

```

This section asks you to build a recipe for the Subaru data to process 
some of those columns. You'll `prep` the recipe on `d.subaru`, then
`juice` it to get the transformed tibble out. Then I've asked you to answer 
some questions about the results. 

### Auto Models

The auto models in this data set are messy. There are 
`r length(unique(d.subaru$model))` different makes and 
`r scales::percent(mean(is.na(d.subaru$model)))` of the values 
are missing. Here's a messy picture: 

```{r}
 d.subaru %>% 
  count(model) %>% 
  mutate(model = fct_reorder(factor(model),.x=n,.fun=sum)) %>% 
  ggplot(aes(x=n,y=model)) + 
  geom_point() + 
  theme_minimal() + 
  labs(x="Number of Rows (Log)",y="Raw Subaru Model") + 
  scale_x_log10(label=scales::comma_format())

```

The top five, (Outback, Forester, Legacy, Impreza, WRX) have enough data to
support modeling. We'd like to map others to an "Other" category. When you build
your recipe, include a step that transforms the `model` column to do this mapping.

### Odometer Data

The odometer data is hand-entered, which makes it messy. We could realistically
expect a log-relationship with mileage, since going from 10K to 20K miles is 
different than going from 200K to 210K. Add a step to your recipe to do 
a log-transform of this column. 

### Missing Data in Year

The `year` column clearly has some incorrect data in it. Additionally, 
the `year` column is completed for `r scales::percent(mean(!is.na(d.subaru$year)))`
of our rows. Using either the K-nearest neighbors or linear model imputation 
methods, add a recipe step that imputes `year`. Pay attention to what covariates
you're using to do the imputation and be prepared to justify your choices. 

## Your Recipe

Use the code block below to add steps to your recipe. 

```{r}

d.prep <- d.subaru %>% 
  recipe() %>% 
  # No steps yet, but yours go here. 
  prep() %>% 
  juice()


```


## Top 5
``` {r recipe}
count_top_five <- d.subaru %>%
group_by(model) %>%
filter(!is.na(model)) %>%
count(model)

count_top_five <- count_top_five %>%
arrange(desc(n))

count_top_five_num <- as.integer(count_top_five[5,2])

d.prep <- d.subaru %>% 
  recipe() %>% 
  step_other(model,threshold = count_top_five_num, other = NA) %>% 
  prep() %>% 
  juice()


d.prep <- d.subaru %>% 
recipe(price~.) %>% 
step_knnimpute(year,
  role = NA,
  trained = FALSE,
  neighbors = 5,
  impute_with = imp_vars(all_predictors()),
  options = list(nthread = 1, eps = 1e-08),
  ref_data = NULL,
  columns = NULL,
  skip = FALSE,
  id = rand_id("knnimpute")
) %>% 
 prep() %>% 
  juice()
```

``` {r log}
d.prep <- d.subaru %>% 
  recipe() %>% 
  step_log(odometer) %>% 
  prep() %>% 
  juice()
```

``` {r plot}

## Plot Original Data
 d.subaru %>% 
  count(year) %>% 
  mutate(year = fct_reorder(factor(year),.x=n,.fun=sum)) %>% 
  ggplot(aes(x=n,y=year))+ 
  geom_point() + 
  theme_minimal() + 
  labs(x="Number of Rows (Log)",y="Raw Subaru Year") + 
  scale_x_log10(label=scales::comma_format())

## Plot Prep Data
 d.prep %>% 
  count(year) %>% 
  mutate(year = fct_reorder(factor(year),.x=n,.fun=sum)) %>% 
  ggplot(aes(x=n,y=year))+ 
  geom_point() + 
  theme_minimal() + 
  labs(x="Number of Rows (Log)",y="Raw Subaru Year") + 
  scale_x_log10(label=scales::comma_format())
```

## Additional Tasks

With your `recipe` written, now some tasks for you: 

1. Re-do the above plot of models on your prepped data. Interpret the results. 
1. Build a plot of odometer reading in the original data versus the transformed
data. Which observations look odd? Were any new infinite values created when you 
took the `log`? How come? What is a quick change to the recipe step that would 
eliminate these? 
1. Make a scatterplot of counts by year in the original data versus in the 
imputed data. Which years show the biggest changes? 

The log number one occur to be zero, and some are way to high for the odeometer to be true.
2004 quickly shifted to the number 4 from 7 on the scatterplot. 96 and 93 swtiched spot on the plot. The different number of row occurs to be smaller in the top and bottom. 

### New Model Data
``` {r new model}
 d.prep %>% 
  count(model) %>% 
  mutate(model = fct_reorder(factor(model),.x=n,.fun=sum)) %>% 
  ggplot(aes(x=n,y=model)) + 
  geom_point() + 
  theme_minimal() + 
  labs(x="Number of Rows (Log)",y="Raw Subaru Model") + 
  scale_x_log10(label=scales::comma_format())

```


## Build Holiday Ranges

The list found at `timeDate::listHolidays()` may be useful for this exercise. 
Using a mix of `recipes` and `dplyr`, add the following columns to our `d.dates`
data set: 

* Add dummy variables for the major holidays in the US: Presidents' Day, 
  MLK's Birthday, Christmas Day, Election Day, New Year's Day, Thanksgiving, 
  Veteran's Day, Independence Day, Memorial Day, Labor Day, Christmas Eve, and
  New Year's Eve. 
  
## use a for loop instead of a step function
* Add a column called `dt_spring` that has a 1 for every day between 
  Easter Sunday and Labor Day. Note that this isn't a canned `step_` function.
  (I had to do that for a model once because this was a busy time for shopping
  for spring clothing lines.)
* Have all your columns begin with the string "dt_" so that we can see them on the 
  plot
  
To get you've started and to test the plotting, I've added one holiday for you. 

```{r}
d.dates <- d.dates %>%
   # mutate(dt_spring = case_when(between(dt, as.Date("Easter"(2018)),as.Date("USMemorialDay"(2018))) ~ 1,
    #                             between(dt, as.Date("Easter"(2019)),as.Date("USMemorialDay"(2019))) ~ 1,
     #                            between(dt, as.Date("Easter"(2020)),as.Date("USMemorialDay"(2020))) ~ 1))
 mutate(dt_spring=0)



  

##Prep  
prepped.date.data <- d.dates %>% 
  recipe() %>% 
  step_holiday(dt, holidays = c("USPresidentsDay","USMLKingsBirthday","ChristmasDay","USElectionDay", "NewYearsDay","USThanksgivingDay","USVeteransDay","USIndependenceDay","USMemorialDay","USLaborDay","ChristmasEve", "DENewYearsEve","Easter" )) %>% 
  prep() %>% 
  juice() %>% 
  mutate(dt_spring=0)

for(i in 1:nrow(prepped.date.data)){
  this.easter <- prepped.date.data$dt_Easter[i]
  this.memorial <- prepped.date.data$dt_USMemorialDay[i]
  
  if(this.easter==1){
    prepped.date.data$dt_spring[i]==1
  }
  

}



timeDate::listHolidays()

prepped.date.data %>% 
  select(starts_with("dt")) %>% 
  pivot_longer(cols=-dt,
               names_to="holiday") %>% 
  ggplot(aes(x=dt,y=value,color=holiday)) + 
  geom_line() + 
  facet_wrap(~holiday) + 
  theme_minimal() + 
  labs(x="Date",y="Holiday Indicator",color="Column")



 

```


## Appendix: Subaru Data Description

I haven't written this yet! It's on my to-do list, though. In the meantime,
ask questions about the data on Teams!


