---
title: "Regression Modeling"
author: "Xuhao Dong"
date: "4/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(recipes)
library(dplyr)
library(tidymodels)
library(here)
library(rpart.plot)
library(xgboost)
d <- read_tsv("subaru_data.txt")
d1 <- d %>% 
    mutate(price=as.numeric(price),
        model = as.factor(model),
        odometer = as.numeric(odometer))
```

```{r cleanup, include=FALSE}
summary(d1$model)
is.na(d1$year) <- d1$year > 2017
is.na(d1$odometer) <- d1$odometer > 300000
is.na(d1$odometer)<- d1$odometer < 100
is.na(d1$price) <- d1$price >40000
is.na(d1$price) <- d1$price <500

is.na(d1$model) <- d1$model=="baja-outback"
is.na(d1$model) <- d1$model=="brz-sti "
is.na(d1$model) <- d1$model=="forester-impreza-legacy-outback-wrx-xt "
is.na(d1$model) <- d1$model=="forester-outback  "
is.na(d1$model) <- d1$model=="forester-sti"
is.na(d1$model) <- d1$model=="forester-tribeca"
is.na(d1$model) <- d1$model=="forester-wrx-xt"
is.na(d1$model) <- d1$model=="impreza-legacy-wrx"
is.na(d1$model) <- d1$model=="impreza-outback-sti"
is.na(d1$model) <- d1$model=="legacy-outback"
is.na(d1$model) <- d1$model=="legacy-sti"
is.na(d1$model) <- d1$model=="legacy-wrx"
is.na(d1$model) <- d1$model=="baja-forester"
is.na(d1$model) <- d1$model=="brz-sti-wrx"
is.na(d1$model) <- d1$model=="forester-legacy-outback"
is.na(d1$model) <- d1$model=="forester-outback-tribeca"
is.na(d1$model) <- d1$model=="forester-sti-wrx"
is.na(d1$model) <- d1$model=="forester-wrx"
is.na(d1$model) <- d1$model=="impreza-legacy"
is.na(d1$model) <- d1$model=="impreza-sti"
is.na(d1$model) <- d1$model=="impreza-wrx"
is.na(d1$model) <- d1$model=="impreza-legacy"
is.na(d1$model) <- d1$model=="legacy-outback-xt"
is.na(d1$model) <- d1$model=="legacy-sti-xt"
is.na(d1$model) <- d1$model=="legacy-xt"
```

## Location

```{r location}
d1 %>%
  group_by(price) %>% 
  filter(!is.na(price))

d1 %>% 
    mutate(crumb_area = fct_reorder(crumb_area, price, .fun='median' )) %>%
    ggplot( aes(x=crumb_area, y=price, fill=crumb_area)) + 
    geom_boxplot() +
    xlab("class") +
    theme(legend.position="none") +
    xlab("") +
    xlab("")


```
Result: Minneapolis has the lowest overall price (lowest 1st, and 3rd quartile). Missoula and Atlanta has similar overall price, with Atlanta has a slightly higher 3rd quartile, and median. 

## Linear Regression

```{r train, echo=FALSE}
d1 %>% 
  filter(!is.na(model)) %>% 
  filter(!is.na(odometer)) %>% 
  filter(!is.na(year))

set.seed(27341)

split <- initial_split(d1)
d.train <- training(split)
d.test <- testing(split)

```


```{r model1, echo=FALSE}
model1 <- linear_reg() %>% 
  set_engine("lm")

fit1 <- model1 %>% 
  fit(price ~ year + odometer, d.train)

tidy(fit1)

```

```{r model2, echo=FALSE}
model2 <- linear_reg() %>% 
  set_engine("lm")

fit2 <- model2 %>% 
  fit(price ~ crumb_area + model, d.train)

tidy(fit2)

```

```{r model3, echo=FALSE}
model3 <- linear_reg() %>% 
  set_engine("lm")

fit3 <- model3 %>% 
  fit(price ~ year + odometer + crumb_area + model, d.train)

tidy(fit3)

```
Result: Model 3 with all the variables in it has the best fit. Year, and car model are the two most signifacnt to the overall car price, with year and location following last. 

``` {r predict, echo=FALSE}
model.tree <- decision_tree() %>% 
  set_engine("rpart") %>% 
  set_mode("regression")

fit.tree <- model.tree %>% 
  fit(price ~ year + odometer + crumb_area + model, d.test)

d.test <- d.test %>% 
  bind_cols(predict(fit.tree, d.test)) %>% 
  rename(pred_tree = .pred)

metrics(d.test,truth =price,estimate = pred_tree)

rpart.plot(fit.tree$fit)


```


``` {r residuals(), echo=FALSE}

fit_all <- boost_tree() %>% 
  set_engine("xgboost") %>% 
  set_mode("regression") %>% 
  fit(price ~ year + odometer + crumb_area + model, d)

#fit.all <- lm(price ~ year + odometer + crumb_area + model, d)

par(mfrow = c(2, 2))  
plot(fit_all) 
res <- residuals(fit_all)
tail(res)
pred<-predict(fit_all)


```




